---
title: "rosetta"
author: "Brett Klamer"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`rosetta` allows an analyst to combine datasets that measure the same latent traits when there is only partial overlap of measurements across the constituent datasets.

# Simulate data

Consider the case where we have three independent datasets which have measurements on three latent factors. In total, we have three variables per latent factor, however, each dataset only measures two out of three per latent factor.

```{r}
library(rosetta)
d_sim <- sim(seed = 100)
d_missing <- d_sim$missing
d_complete <- d_sim$complete
```

The simulated 'complete' data would look like

```{r}
lapply(d_complete, head)
```

while the simulated 'missing' data (representative of our real life use case) looks like

```{r}
lapply(d_missing, head)
```

# Run rosetta

'rosetta' can now be run so that the latent factors contained within the three independent datasets are summarized into a single dataset of factor scores. This allows simplicity, statistical power, and modeling flexibility of a single joint analysis of the information contained within the original data.

```{r}
d_rosetta <- rosetta(
  d = d_missing,
  factor_structure = list(
    a = c("a_1", "a_2", "a_3"),
    b = c("b_1", "b_2", "b_3"),
    c = c("c_1", "c_2", "c_3")
  )
)

# combine rosetta results into a single dataset
d_rosetta <- as.data.frame(do.call("rbind", d_rosetta))

# check the factor score output
head(d_rosetta)
```

# Compare results

We will compare the factor scores from the complete data versus the missing data.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# get factor scores from complete data
## bind the complete data
d_complete <- do.call("rbind", d_complete)
## create RAM model
factor_structure <- attributes(d_sim)$factor_structure
sem_model <- rosetta:::sem_model(factor_structure)
## observed covariance matrix
cov_mat <- rosetta:::obs_cov(d_complete)
## complete sem
sem_fit <- sem::sem(
  model = sem_model,
  S = cov_mat,
  N = ncol(cov_mat)
)
## model results
complete_fscores <- as.data.frame(sem::fscores(model = sem_fit, data = d_complete))

# Visualize comparison
## combine data
d_rosetta <- tidyr::gather(d_rosetta, key = "key", value = "rosetta", a, b, c)
complete_fscores <- tidyr::gather(complete_fscores, key = "key", value = "complete", a, b, c)
d_plot <- cbind(d_rosetta, complete_fscores["complete"])
## plot
ggplot(d_plot, aes(x = complete, y = rosetta, color = key)) + 
  geom_point() + 
  labs(
    x = "Complete data factor scores",
    y = "Rosetta factor scores",
    color = "Factor"
  )
## correlation table
table <- d_plot %>%
  dplyr::group_by(key) %>%
  dplyr::summarize(cor = cor(rosetta, complete, method = "pearson"))
knitr::kable(table, caption = "correlation table")
```
