---
title: "rosetta"
author: "Brett Klamer (updated by Brendan Galdo)"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`rosetta` allows an analyst to combine datasets that measure the same latent traits when there is only partial overlap of measurements across the constituent datasets.

# Simulate data

Consider the case where we have three independent datasets which have measurements on three latent factors. In total, we have three variables per latent factor, however, each dataset only measures two out of three per latent factor.

```{r,message=FALSE}
# 
library(devtools)
load_all()

#library(rosetta)
library(dplyr)
library(tidyr)
library(ggplot2)


```

```{r}

d_sim <- sim(seed = 100)
d_missing <- d_sim$missing
d_complete <- d_sim$complete

```

The simulated 'complete' data would look like

```{r}

lapply(d_complete, head)

```

while the simulated 'missing' data (representative of our real life use case) looks like

```{r}

lapply(d_complete, head)
```

# Run rosetta

'rosetta' can now be run so that the latent factors contained within the three independent datasets are summarized into a single dataset of factor scores. This allows simplicity, statistical power, and modeling flexibility of a single joint analysis of the information contained within the original data.

```{r}
d_rosetta <- rosetta(
  d = d_missing,
  factor_structure = list(
    a = c("a_1", "a_2", "a_3"),
    b = c("b_1", "b_2", "b_3"),
    c = c("c_1", "c_2", "c_3")
  ),
  id_colnames = "ID"
)

# combine rosetta results into a single dataset
d_rosetta <- as.data.frame(do.call("rbind", d_rosetta))

# check the factor score output
head(d_rosetta)

```

# Compare results

We will compare the factor scores from the complete data versus the missing data.

```{r}


# get factor scores from complete data 

d_rosetta_complete = rosetta(
  d = d_complete,
  factor_structure = list(
    a = c("a_1", "a_2", "a_3"),
    b = c("b_1", "b_2", "b_3"),
    c = c("c_1", "c_2", "c_3")
  ),
  id_colnames = "ID"
)

# combine rosetta results (for the complete) into a single dataset
d_rosetta_complete <- as.data.frame(do.call("rbind", d_rosetta_complete))

# create a data.frame of results for plotting
plot_df = data.frame("factor_scores_incomplete_data" = c(d_rosetta$a,
                                                         d_rosetta$b,
                                                         d_rosetta$c),
                     "factor_scores_complete_data" = c(d_rosetta_complete$a,
                                                       d_rosetta_complete$b,
                                                       d_rosetta_complete$c),
                     "latent_variable" = as.factor(c(rep("a",length(d_rosetta_complete$a)), rep("b",length(d_rosetta_complete$a)),rep("c",length(d_rosetta_complete$a)))))

ggplot(plot_df) + geom_point(aes(x = factor_scores_incomplete_data,
                               y = factor_scores_complete_data))+facet_grid(rows="latent_variable")

```

